Concurrent Programming with CXL

### 2020-06-17 book/paper.tex
Latest

We suggested the last technique to address the deadlock issue in the Dining
Philosophers problem (Chapter~\ref{ch:deadlock}).

@@

Figure~\ref{fig:dinerscv} demonstrates how this might be done.
We use a single mutex for the diners, and, for each fork, a boolean
and a condition variable.  The boolean indicates if the fork has been
taken.
Each diner waits if either the left or right fork is already taken.
But which condition variable to wait on?
The code demonstrates an important technique to use when waiting for multiple
conditions.
The condition in the \texttt{while} statement is the condition that the
diner is waiting for to be \texttt{False}.  Within the \texttt{while} statement,
there is an \texttt{if} statement for each disjunct of that condition.
The code waits for either or both forks if necessary.  After that, it goes
back to the top of the \texttt{while} loop.

@@

\begin{quote}
\begin{itemize}
\item Can you see why this does not work?  What can go wrong?
\end{itemize}
\end{quote}

The next approach suggested by Havender is to allow processes to back out.
While this could be done, this invariably leads to a busy waiting solution
where a process keeps obtaining locks and releasing them again until it
finally is able to get all of them.
The approach to order the resources was already presented in Chapter~\ref{ch:deadlock}

There is, however, another approach. which is sometimes called
\emph{deadlock avoidance}.
\index{deadlock avoidance}.

>>>>

@@

Figure~\ref{fig:dinerscv} demonstrates how this might be done.
We use a single mutex for the diners, and, for each fork, a boolean
and a condition variable.  The boolean indicates if the fork has been
taken.
Each diner waits if either the left or right fork is already taken.
But which condition variable to wait on?
The code demonstrates an important technique to use when waiting for
multiple conditions.
\index{multiple conditions, waiting on}
The condition in the \texttt{while} statement is the condition that the
diner is waiting for to be \texttt{False}.  Within the \texttt{while} statement,
there is an \texttt{if} statement for each disjunct of that condition.
The code waits for either or both forks if necessary.  After that, it goes
back to the top of the \texttt{while} loop.

@@

\begin{quote}
\begin{itemize}
\item Can you see why this does not work?  What can go wrong?
\item Run it through CXL in case you are not sure!
\end{itemize}
\end{quote}

The \texttt{Preemption} approach suggested by Havender is to allow processes to back out.
While this could be done, this invariably leads to a busy waiting solution
where a process keeps obtaining locks and releasing them again until it
finally is able to get all of them.
The \emph{No Circular Waiting} approach to order the resources was already presented in
Chapter~\ref{ch:deadlock}

There is, however, another approach. which is sometimes called
\emph{deadlock avoidance}.
\index{deadlock avoidance}.
In the case of the Dining Philosophers, we want to avoid the situation where each
diner picks up a fork.  So if we can prevent more than four diners from starting to
eat at the same time, then we can avoid the conditions for deadlock from ever
happening.
Figure~\ref{fig:dinersavoid} demonstrates this concept.  It uses a semaphore to
restrict the number of diners at any time to four.

This concept can be generalized using something called the
Banker's Algorithm~\cite{EWD108}, but it is outside the scope of this book.
The problem with these kinds of schemes is that one needs to know ahead of time
what the maximum number of resources is that each process wants to allocate,
making them generally quite impractical.
